
## Simulate non-spatial, closed population cap-recap data: M0


Parameters

```{r}
N <- 30
K <- 10  # nOccasions
p <- 0.1 # capture prob on 1 occasion

1 - (1-p)^K # capture prob after K occasions
```


All (observed and unobserved) capture histories

```{r}
yall <- matrix(NA, N, K)
for(k in 1:K) {
    yall[,k] <- rbinom(N, 1, p)
}
```


Discard unobservable (all zero) capture histories

```{r}
y <- yall[rowSums(yall)>0,]
y
```



Summary stats

```{r}
rowSums(y)
table(rowSums(y))
colSums(y)
```



# Bayesian analysis

## Joint posterior

$$
p(z,\psi,p|y) \propto p(y|z,p)p(z|\psi)p(\psi)p(p)
$$



## Gibbs sampler in R

```{r}

M0.gibbs <- function(y, M, niter, tune) {

    n0 <- nrow(y)
    K <- ncol(y)
    yaug <- matrix(0L, M, K)
    yaug[1:n0,] <- y
    captured <- rowSums(yaug, na.rm=TRUE)>0

    ## Initial values
    z <- rep(1, M)
    psi <- runif(1)
    p <- runif(1)

    ## Initial densities
    ## Avoiding for loop with R's recycling rules
    lp.y <- dbinom(yaug, 1, z*p, log=TRUE)
    lp.z <- dbinom(z, 1, psi, log=TRUE)
    lp.psi <- dbeta(psi, 1, 1, log=TRUE)
    lp.p <- dbeta(p, 1, 1, log=TRUE)

    out <- matrix(NA, niter, 4)
    colnames(out) <- c("N", "psi", "p", "deviance")

    for(iter in 1:niter) {
        ## Sample from p(z|dot)
        for(i in 1:M) {
            if(captured[i])
                next
            z.cand <- 1-z[i]
            lp.z.cand <- dbinom(z.cand, 1, psi, log=TRUE)
            lp.y.cand <- dbinom(yaug[i,], 1, z.cand*p, log=TRUE)
            if(runif(1) < exp((lp.z.cand+sum(lp.y.cand)) -
                              (lp.z[i]+sum(lp.y[i,])))) {
                z[i] <- z.cand
                lp.y[i,] <- lp.y.cand
                lp.z[i] <- lp.z.cand
            }
        }

        ## Sample from p(psi|dot)
        psi.cand <- rnorm(1, psi, tune[1])
        if(psi.cand>0 & psi.cand<1) {
            lp.z.cand <- dbinom(z, 1, psi.cand, log=TRUE)
            lp.psi.cand <- dbeta(psi.cand, 1, 1, log=TRUE)
            if(runif(1) < exp((sum(lp.z.cand)+lp.psi.cand)-
                              (sum(lp.z)+lp.psi))) {
                psi <- psi.cand
                lp.z <- lp.z.cand
                lp.psi <- lp.psi.cand
            }
        }
        ## psi <- rbeta(1, 1+sum(z), 1+M-sum(z))

        ## Sample from p(p|dot)
        p.cand <- rnorm(1, p, tune[2])
        if(p.cand>0 & p.cand<1) {
            lp.y.cand <- dbinom(yaug, 1, z*p.cand, log = TRUE)
            lp.p.cand <- dbeta(p.cand, 1, 1, log = TRUE)
            if(runif(1) < exp((sum(lp.y.cand)+lp.p.cand)-
                              (sum(lp.y)+lp.p))) {
                p <- p.cand
                lp.y <- lp.y.cand
                lp.p <- lp.p.cand
            }
        }

        N <- sum(z)
        deviance <- -2*sum(lp.y)

        out[iter,] <- c(N, psi, p, deviance)

    }
    return(out)
}

```



```{r}

## debugonce(M0.gibbs)

fmM0.1 <- M0.gibbs(y=y, M=100, niter=1000, tune=c(0.15,0.05))

library(coda)
mcM0.1 <- as.mcmc(fmM0.1)
plot(mcM0.1)

rejectionRate(mcM0.1)


```









## JAGS


Augment the data

```{r}
M <- 100
yAug <- matrix(0, M, K)
yAug[1:nrow(y),] <- y
##yAug
```

library(rjags)

## Model M0
jd <- list(yAug=yAug, K=K, M=M)

ji <- function() list(p=runif(1), psi=runif(1),
                      z=rep(1, M))

jp <- c("p", "psi", "N", "deviance")

load.module("dic")

jm <- jags.model(file="M0.jag", data=jd, inits=ji)

jp <- coda.samples(jm, jp, n.iter=1000)


plot(jp)

summary(jp)






## Model Mt
jd <- list(yAug=yAug, K=K, M=M)

ji.Mt <- function() list(p=runif(jd$K), psi=runif(1),
                      z=rep(1, M))
ji.Mt()

jp <- c("p", "psi", "N", "deviance")


jm.Mt <- jags.model(file="Mt.jag", data=jd, inits=ji.Mt)

jp.Mt <- coda.samples(jm.Mt, jp, n.iter=1000)


plot(jp.Mt, ask=TRUE)

(sp.Mt <- summary(jp.Mt))

plot(1:10, sp.Mt$quantile[2:11,3], ylim=c(0,1))
segments(1:10, sp.Mt$quantile[2:11,1], 1:10, sp.Mt$quantile[2:11,5])









## Model Mb

firstcap <- apply(y, 1, function(x) min(which(x==1)))
firstcap

prevcap <- matrix(1, M, ncol(y))
for(i in 1:nrow(y)) {
    prevcap[i, (firstcap[i]+1):ncol(y)] <- 2
}

prevcap

jd.Mb <- list(yAug=yAug, K=K, M=M, prevCap=prevcap)

ji.Mb <- function() list(p=runif(2), psi=runif(1),
                         z=rep(1, M))
ji.Mb()

jp <- c("p", "psi", "N", "deviance")


jm.Mb <- jags.model(file="Mb.jag", data=jd.Mb, inits=ji.Mb)

jp.Mb <- coda.samples(jm.Mb, jp, n.iter=1000)


plot(jp.Mb, ask=TRUE)

(sp.Mb <- summary(jp.Mb))

plot(1:2, sp.Mb$quantile[3:4,3], ylim=c(0,1), xlim=c(0,3))
segments(1:2, sp.Mb$quantile[3:4,1], 1:2, sp.Mb$quantile[3:4,5])
